### 2017-03-25 paper/basicSkeletons.tex
added |>>>| as a combinator for pipe

\subsection{parEvalNLazy}
\begin{center}
	\includegraphics[scale=0.7]{images/parEvalNLazy}
\end{center}
\code{parEvalN} is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. \code{map (arr . (+)) [1..]} or just because we need the arrows evaluated in chunks. \code{parEvalNLazy} fixes this. It works by first chunking the input from \code{[a]} to \code{[[a]]} with the given \code{ChunkSize} in \code{(arr \$ chunksOf chunkSize)}. These chunks are then fed into a list \code{[arr [a] [b]]} of parallel arrows created by feeding chunks of the passed \code{ChunkSize} into the regular parEvalN by using \code{listApp}. The resulting \code{[[b]]} is lastly converted into \code{[b]} with \code{arr concat}.
\begin{lstlisting}[frame=htrbl]
parEvalNLazy :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> [arr a b] -> (arr [a] [b])
parEvalNLazy conf chunkSize fs =
	arr (chunksOf chunkSize) >>>
	listApp fchunks >>>
	arr concat
	where
		fchunks = map (parEvalN conf) $ chunksOf chunkSize fs
\end{lstlisting}

\subsection{parEval2}
\begin{center}
	\includegraphics[scale=0.7]{images/parEval2}
\end{center}
We have only talked about the paralellization arrows of the same type until now. But sometimes we want to paralellize heterogenous types as well. However, we can implement such a \code{parEval2} combinator which combines two arrows \code{arr a b} and \code{arr c d} into a new parallel arrow \code{arr (a, c) (b, d)} quite easily with the help of the \code{ArrowChoice} typeclass. The idea is to use the \code{+++} combinator which combines two arrows \code{arr a b} and \code{arr c d} and transforms them into \code{arr (Either a c) (Either b d)} to get a common arrow type that we can then feed into parEvalN.
\\\\
We start by transforming the \code{(a, c)} input into a 2-element list \code{[Either a c]}. Then, we feed this list into a parallel arrow constructed with \code{parEvalN} that uses \code{f +++ g}  we then feed into a 2-element parallel arrow that we created as described above. After the calculation is finished we convert the resulting \code{[Either b d]} into \code{([b] , [d])}. The two lists in this tuple contain only 1 element each by construction, so we can finally just convert the tuple to \code{(b, d)} in the last step.
\begin{lstlisting}[frame=htrbl]
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g = 
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{lstlisting}

>>>>

\subsection{parEvalNLazy}
\begin{center}
	\includegraphics[scale=0.7]{images/parEvalNLazy}
\end{center}
\code{parEvalN} is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. \code{map (arr . (+)) [1..]} or just because we need the arrows evaluated in chunks. \code{parEvalNLazy} fixes this. It works by first chunking the input from \code{[a]} to \code{[[a]]} with the given \code{ChunkSize} in \code{arr (chunksOf chunkSize)}. These chunks are then fed into a list \code{[arr [a] [b]]} of parallel arrows created by feeding chunks of the passed \code{ChunkSize} into the regular parEvalN by using \code{listApp}. The resulting \code{[[b]]} is lastly converted into \code{[b]} with \code{arr concat}.
\begin{lstlisting}[frame=htrbl]
parEvalNLazy :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> [arr a b] -> (arr [a] [b])
parEvalNLazy conf chunkSize fs =
	arr (chunksOf chunkSize) >>>
	listApp fchunks >>>
	arr concat
	where
		fchunks = map (parEvalN conf) $ chunksOf chunkSize fs
\end{lstlisting}

\subsection{parEval2}
\begin{center}
	\includegraphics[scale=0.7]{images/parEval2}
\end{center}
We have only talked about the paralellization arrows of the same type until now. But sometimes we want to paralellize heterogenous types as well. However, we can implement such a \code{parEval2} combinator which combines two arrows \code{arr a b} and \code{arr c d} into a new parallel arrow \code{arr (a, c) (b, d)} quite easily with the help of the \code{ArrowChoice} typeclass. The idea is to use the \code{+++} combinator which combines two arrows \code{arr a b} and \code{arr c d} and transforms them into \code{arr (Either a c) (Either b d)} to get a common arrow type that we can then feed into parEvalN.
\\\\
We start by transforming the \code{(a, c)} input into a 2-element list \code{[Either a c]} by first tagging the two inputs with \code{Left} and \code{Right} and wrapping the right element in a singleton list with \code{return} so that we can combine them with \code{arr (uncurry (:))}. Next, we feed this list into a parallel arrow running on 2 instances of \code{f +++ g} as described above. After the calculation is finished we convert the resulting \code{[Either b d]} into \code{([b], [d])} with \code{arr partitionEithers}. The two lists in this tuple contain only 1 element each by construction, so we can finally just convert the tuple to \code{(b, d)} in the last step.
\begin{lstlisting}[frame=htrbl]
parEval2 :: (ArrowChoice arr,
	ArrowParallel arr (Either a c) (Either b d) conf) =>
	conf -> arr a b -> arr c d -> arr (a, c) (b, d)
parEval2 conf f g = 
	arr Left *** (arr Right >>> arr return) >>>
	arr (uncurry (:)) >>>
	parEvalN conf (replicate 2 (f +++ g)) >>>
	arr partitionEithers >>>
	arr head *** arr head
\end{lstlisting}

### 2017-03-25 paper/conclusion.tex
added |>>>| as a combinator for pipe

\section{Conclusion}
As we have seen in this paper, arrows are a useful tool for composing parallel programs. By basing our parallel Haskell on them, we do not have to introduce new monadic types that wrap the computation and instead can use them just like they were regular sequential pure code. Performancewise, parallel arrows are on par with existing parallel Haskells, as they do not introduce any notable overhead. While we have seen that the parallel arrows can express computation on clusters, it is noteworthy that in its current state, the parallel arrow API doesn't have as much control over data-flow as for example Eden. In its current state this will end up in the root node being a possible bottleneck on clusters when parallel arrows are combined with \code{>>>}. This can probably be remedied by introducing some sort of a Java-like Future construct that can be passed between nodes so that nodes can communicate without going through the root node. This will have to be implemented in a follow up paper, though.

>>>>

\section{Conclusion}
As we have seen in this paper, arrows are a useful tool for composing parallel programs. By basing our parallel Haskell on them, we do not have to introduce new monadic types that wrap the computation and instead can use them just like they were regular sequential pure code. Performancewise, parallel arrows are on par with existing parallel Haskells, as they do not introduce any notable overhead.

\fixme{ArrowApply (or equivalent) are needed because we basically want to be able to produce intermediary results, this is by definition of the parallel evaluation combinators}

### 2017-03-25 paper/topologySkeletons.tex
added |>>>| as a combinator for pipe

>>>>

\begin{lstlisting}[frame=htrbl]
(|>>>|) :: (ArrowLoop arr, ArrowChoice arr,
	ArrowParallel arr (fut (([a], [b]), [c])) (fut (([a], [b]), [c])) (),
	Future fut (([a], [b]), [c])) =>
	arr a b -> arr b c -> arr a c
	(|>>>|) f g =
		arr (\a -> (([a], []), [])) >>>
		pipe () (replicate 2 $ unify f g) >>> arr snd >>> arr head
			where
				unify :: (ArrowChoice arr) =>
					arr a b -> arr b c -> arr (([a], [b]), [c]) (([a], [b]), [c])
				unify f g =
					(mapArr f *** mapArr g) *** arr (\_ -> []) >>>
					arr (\((a, b), c) -> ((c, a), b))
\end{lstlisting}
